---
title: "average wage select counties"
output: github_document
---


```{r, message=FALSE}
library(tidyverse)
```

```{r}
load(here::here("data", "acs2019.RData"))
```

joining fips codes into one code so we can use the codes given by fact finder
```{r}
counties <- acs2019 %>% 
  unite("state_county", statefip, countyfip, sep = "", na.rm = TRUE, remove = FALSE)
```

filtering counties of interest
```{r}
counties <- counties %>% 
  filter(state_county %in% c(2681, 39153, 183, 1743, 5525))

counties$state_county <- factor(counties$state_county, labels = c("Dupage", "Allen", "Kent", "Summit", "Dane"))

counties <- counties %>% 
  mutate(highest_deg = case_when(educd < 62 ~ "Less than HS",
                                 educd < 81 ~ "High School",
                                 educd < 101 ~ "Associates",
                                 educd < 114 ~ "Bachelors",
                                 TRUE ~ "Grad"),
         highest_deg = factor(highest_deg)) #will need to reconsider these blocks

counties <- counties %>% 
  rename(county = state_county)
```

As a reminder, table below shows what city a county is a proxy for

|County Proxy|City|
|------------|----|
|Dupage      |Naperville, IL|
|Allen       |Fort Wayne, IN|
|Kent        |Grand Rapids, MI|
|Summit      |Akron, OH|
|Dane        |Madison, WI|


summary table. The data is so skewed right this probably isn't very useful. Maybe do weighted median?
```{r}
counties %>% 
  group_by(county) %>% 
  summarize(mean_wage = weighted.mean(incwage, perwt), n = n()) %>% 
  knitr::kable()
```

graphing
```{r}
#when am I supposed to use weights
counties %>% 
  ggplot(aes(x = incwage))+
  geom_boxplot(aes(weight = perwt, color = county))+
  labs(x = "Income",
       color = "County",
       title = "Income by county",
       subtitle = "weighted?")+
  scale_x_continuous(labels = scales::dollar,
                     expand = expansion(mult = c(.01, .06)))+
  ggpubr::theme_pubr()+
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank())
```

ANOVA to see if there is a significant difference between the counties
Normality condition technically violated but we can maybe rely on CLT? Could also remove outliers or focus on IQR. Or use KW test.
There are also unequal variances, unsure if largest variance is more than 4 times the smallest. Regardless, Welch's Anova from Rstatix may be the way to go.
```{r}
anova <- aov(incwage ~ county, weights = perwt, data = counties)
summary(anova)
```
We conclude from the ANOVA that there is a significant difference in earnings between the counties. We can use Tukey to find out which differences specifically are different. (Could use Bonferroni to be more conservative)
```{r}
TukeyHSD(anova)

```
The significant differences are

- Summit and Dupage
- Summit and Allen
- Dane and Allen
- Summit and Kent

Looking closely at the Tukey output, the outliers are really skewing the results. It may be beneficial to either remove outliers or look only within IQR.

The cities may also not have been chosen well to be compared to Kent County (which is what we're really interested in)



## Just looking at people with Bachelors

```{r}
undergrad <- counties %>% 
  filter(highest_deg == "Bachelors")

undergrad %>% 
  ggplot(aes(x = incwage))+
  geom_boxplot(aes(weight = perwt, color = county))+
  labs(x = "Income",
       color = "County",
       title = "Income by county",
       subtitle = "weighted? Bachelor's degree only")+
  scale_x_continuous(labels = scales::dollar,
                     expand = expansion(mult = c(.01, .06)))+
  ggpubr::theme_pubr()+
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank())
```

Normality and variance assumptions likely violated. KW or Welch anova may be best. but I'll continue with ANOVA for now.

```{r}
under_anova <- aov(incwage ~ county, weights = perwt, data = undergrad)
summary(under_anova)
```
significant difference between counties tukey to do pair-wise

```{r}
TukeyHSD(under_anova)
```
Significant at 5%

- Allen Dupage (dupage pays more)
- Kent Dupage (dupage pays more)
- Dane Dupage (dupage pays more)

## Conclusion

At least with the counties chosen, Grand Rapids doesn't pay significantly less than most of the counties. The counties chosen were intended to be similar to Grand Rapids but there may be value in including larger cities nearby like Detroit or Chicago. It may also be worthwhile to include counties outside of the midwest as difference in pay may be more of a regional thing.

We should also make age-earnings profiles for the counties to see if there's a difference on age. We could do that with ANOVA but that sounds painful... Maybe with some form of factor (early/mid/late career).

We'll also want to break it down by majors
