---
title: "average wage select counties"
output: github_document
---


```{r, message=FALSE}
library(tidyverse)
```

```{r}
load(here::here("data", "acs2019.RData"))
```

joining fips codes into one code so we can use the codes given by fact finder
```{r}
counties <- acs2019 %>% 
  unite("state_county", statefip, countyfip, sep = "", na.rm = TRUE, remove = FALSE)
```

filtering counties of interest
```{r}
counties <- counties %>% 
  filter(state_county %in% c(2681, 39153, 183, 1743, 5525))

counties$state_county <- factor(counties$state_county, labels = c("Dupage", "Allen", "Kent", "Summit", "Dane"))

counties <- counties %>% 
  rename(county = state_county)
```

summary table :)
```{r}
counties %>% 
  group_by(county) %>% 
  summarize(mean_wage = weighted.mean(incwage, perwt), n = n()) %>% 
  knitr::kable()
```

graphing
```{r}
#when am I supposed to use weights
counties %>% 
  ggplot(aes(x = incwage))+
  geom_boxplot(aes(weight = perwt, color = county))+
  labs(x = "Income",
       color = "County",
       title = "Income by county",
       subtitle = "weighted?")+
  scale_x_continuous(labels = scales::dollar,
                     expand = expansion(mult = c(.01, .06)))+
  ggpubr::theme_pubr()+
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank())
```

ANOVA to see if there is a significant difference between the counties
```{r}
anova <- aov(incwage ~ county, weights = perwt, data = counties)
summary(anova)
```
We conclude from the ANOVA that there is a significant difference in earnings between the counties. We can use Tukey to find out which differences specifcally are different.
```{r}
TukeyHSD(anova)

```
The significant differences are

- Summit and Dupage
- Summit and Allen
- Dane and Allen
- Summit and Kent

