---
title: "general analysis"
author: "Jason Smith"
date: "4/19/2021"
output: github_document
---

```{r, message=FALSE}
library(tidyverse)
library(estimatr)
library(ggpubr)
library(Weighted.Desc.Stat)
library(emmeans)
```

load data
```{r}
load(here::here("data", "acs2019.RData"))
load(here::here("data", "emsi.RData"))
```

get highest degree from ACS
```{r}
acs2019 <- acs2019 %>%
  mutate(highest_deg = case_when(educd < 62 ~ "Less than HS",
                                 educd < 81 ~ "High School",
                                 educd < 101 ~ "Associates",
                                 educd < 114 ~ "Bachelors",
                                 TRUE ~ "Grad"),
         highest_deg = factor(highest_deg),
         incwage = case_when(incwage == 999999 ~ NA_real_,
                             incwage == 999998 ~ NA_real_,
                             TRUE ~ incwage))

acs2019 <- acs2019 %>% 
  unite("state_county", statefip, countyfip, sep = "", na.rm = TRUE, remove = FALSE)
```

only people making less than 200k
```{r}
acs2019 <- acs2019 %>% 
  filter(incwage <= 200000)
```

# Do college grads in GR (kent county) earn more than than those without college
```{r}
kent <- acs2019 %>% 
  filter(state_county == 2681)
```

```{r}
kent %>% 
  group_by(highest_deg) %>% 
  summarise(mean = weighted.mean(incwage, perwt, na.rm = TRUE),
            sd = w.sd(incwage, perwt),
            n = n()) %>% 
  knitr::kable()
```

```{r}
kent %>% 
  ggplot(aes(x = incwage))+
  geom_boxplot(aes(color = highest_deg, weight = perwt))+
  scale_x_continuous(labels = scales::dollar)+
  labs(title = "Income by Educational Attainment: Kent County",
       subtitle = "Income capped at $200,000, weighted",
       x = "Income",
       color = "Highest Degree Earned",
       caption = "Source: 2019 ACS")+
  theme_classic()+
  theme(axis.text.y = element_blank())
```

In Kent county, people with a college degree earn more than people without a college degree. Specifically, people with a Bachelors makes $31,869.43 more on average

compare with EMSI data to look at in-region alumni (bachelors only)
```{r}
emsi <- emsi %>% 
  filter(Location.Status == "Resides in Region" & Level == "Baccaureate") %>% 
  uncount(Freq)
```

```{r}
emsi %>% 
  summarise(mean = mean(Avg..of.Avg..Wage, na.rm = TRUE),
            sd = sd(Avg..of.Avg..Wage),
            n = n()) %>% 
  knitr::kable()
```

```{r}
emsi %>% 
  ggplot(aes(x = Avg..of.Avg..Wage))+
  geom_boxplot()+
  scale_x_continuous(labels = scales::dollar)+
  labs(title = "Average income of GVSU Alumni within Region",
       x = "Average Wage",
       caption = "Source: EMSI, 2020")+  
  theme_classic()+
  theme(axis.text.y = element_blank())
```

GVSU graduates make more on average than college graduates in general by ~$12,425

**the 'in region' defined by EMSI probably includes more than Kent County so the estimate may be off**
**at the very least we can get away with saying GVSU grads don't make less than other college grads in the area**

# major breakdown
```{r}
acs_majors <- acs2019 %>% 
  filter(degfield %in% c(62, 21, 24, 36, 37, 50, 61)) %>% 
  mutate(field = case_when(degfield == 62 ~ "business",
                           degfield %in% c(21, 24, 36, 37, 50) ~ "stem",
                           degfield == 61 ~ "healthcare"),
         field = as.factor(field))
```

```{r}
emsi_majors <- emsi %>%
  mutate(cip_gen = round(CIP.Code)) %>% 
  filter(cip_gen %in% c(52, 11, 14, 26, 27, 40, 51)) %>% 
  mutate(field = case_when(cip_gen == 52 ~ "business",
                           cip_gen %in% c(11, 14, 26, 27, 40) ~ "stem",
                           cip_gen == 51 ~ "healthcare"))

```

```{r}
kent_majors <- acs_majors %>% 
  filter(state_county == 2681)
```

# average wage of Kent by major group
```{r}
kent_majors %>% 
  group_by(field) %>% 
  summarise(mean = weighted.mean(incwage, perwt, na.rm = TRUE),
            sd = w.sd(incwage, perwt),
            n = n()) %>% 
  knitr::kable()
```

# average wage of GV alumni by major group
```{r}
emsi_majors %>% 
  group_by(field) %>% 
  summarise(mean = mean(Avg..of.Avg..Wage),
            sd = sd(Avg..of.Avg..Wage),
            n = n()) %>% 
  knitr::kable()
```

- GV business majors earn about $8,072 more than business majors in Kent County
- GV healthcare majors earn about $7,308 more than healthcare majors in Kent County
- GV STEM majors earn about $6,793 more than STEM majors in Kent County


# Does the GR region pay less than other midwest cities?
```{r}
counties <- acs2019 %>% 
  filter(state_county %in% c(2681, 39153, 183, 1743, 5525) & highest_deg == "Bachelors") %>% 
  mutate(state_county = factor(state_county, labels = c("Dupage", "Allen", "Kent", "Summit", "Dane"))) %>% 
  rename(county = state_county)
```
As a reminder, table below shows what city a county is a proxy for

|County Proxy|City|
|------------|----|
|Dupage      |Naperville, IL|
|Allen       |Fort Wayne, IN|
|Kent        |Grand Rapids, MI|
|Summit      |Akron, OH|
|Dane        |Madison, WI|

```{r}
counties %>% 
  group_by(county) %>% 
  summarize(mean_wage = weighted.mean(incwage, perwt, na.rm = TRUE),
            sd = w.sd(incwage, perwt),
            n = n()) %>% 
  knitr::kable()
```
```{r}
counties %>% 
  ggplot(aes(x = incwage))+
  geom_boxplot(aes(weight = perwt, color = county))+
  scale_x_continuous(labels = scales::dollar)+
  labs(title = "Income by County",
       subtitle = "Income capped at $200,000, weighted",
       x = "Income",
       color = "County",
       caption = "Source: 2019 ACS")+
  theme_classic()+
  theme(axis.text.y = element_blank())
```

using ANOVA, we'll see if there is a significant difference between the earnings of the counties.
The ANOVA assumptions of normality and constant variance aren't really met. We'll rely on CLT for normality. For constant variance, we should use welch's ANOVA but that doesn't let me apply weights and I think the weights are important. ANOVA is fairly robust so hopefully it won't be a big deal. As consolation, we'll use the Holm's method for pairwise comparision since it's more conservative.
```{r}
county_aov <- aov(incwage ~ county, weights = perwt, data = counties)
summary(county_aov)
```
There is a significant difference between the counties, We'll do a post-hoc pairwise comparision with Bonferroni correction to determine which counties are different.

```{r}
summary(pairs(emmeans(county_aov, "county")), adjust = "holm")
```

Of the differences that are significant to the 5%,

- Dupage pays more than Kent
- Dupage pays more than Summit

# Does the pay of college majors differ between counties?
```{r}
counties_majors <- counties %>% 
  filter(degfield %in% c(62, 21, 24, 36, 37, 50, 61)) %>% 
  mutate(field = case_when(degfield == 62 ~ "business",
                           degfield %in% c(21, 24, 36, 37, 50) ~ "stem",
                           degfield == 61 ~ "healthcare"))
```

```{r}
counties_majors_aov <- aov(incwage ~ county*field, weights = perwt, data = counties_majors)
summary(counties_majors_aov)
```

There's not a significant difference in the wages of college graduates between the selected counties. However, there is a significant difference in wages between fields, so we'll do pairwise testing on field.

```{r}
summary(pairs(emmeans(counties_majors_aov, "field")), adjust = "holm")
```

Of the difference significant to the 5%

- Business pays more than healthcare
- STEM pays more than healthcare

Note: the pairwise comparision isn't misleading in this case because all the difference found to be significant are always true among the counties.
```{r}
emmip(counties_majors_aov, field ~ county)
```

In some counties, STEM majors make more than business majors, but the comparision didn't say the difference found between the two majors was significant.


# conclusions

- college graduates make more than people without a college degree in Kent County
- GVSU graduates in the region make more than the average wage of college graduates in Kent County (see note on EMSI's definition of region)
- This difference in earnings between GVSU graduates and other graduates in Kent County is true across all the major groups
- Of the selected counties, Dupage pays more than Kent County, this isn't super surprising since Dupage county is right outside Chicago
- Wages in Kent County was lower than the other counties, but the differences weren't statistically significant
- Healthcare major group made significantly less than both business and STEM
- The differences in wages between the major groups and county weren't statistically significant


# cavets

- We don't know how EMSI defines GVSU's region but it's probably more than Kent county. Possible the more successful alumni in the region live outside Kent County
- The major categories may be slightly different among the EMSI and ACS data. The ACS doesn't use CIP codes for classifying majors
- The assumptions for ANOVA analysis weren't perfectly met, We chose to continue with ANOVA anyways so we could perserve the weighting of the ACS. Using a more conservative p-value adjustment (holms instead of tukey) for pairwise will hopfully account for any errors.
- Comparison counties are all in the Midwest (east north central specifically). Wages may just be lower in this region in general
- ACS doesn't include specific cities so we used counties as a proxy.
- Comparison counties were imperfectly selected. A more careful way of choosing "similar counties" may yield different results.
- Analysis only includes people making less than 200k. I don't included wealthier people would change anything but who knows.